---
title: "Assignment 9: Class Wrap-up"
format: html
editor: visual
author: "Amber Camp"
date: "`r Sys.Date()`"
---

## Assignment 9: Class Wrap-up

We are finally at the end of the semester! The final assignment for the class will be to summarize and share some of the things we've done so far.

The plan is for everyone to be randomly paired with a tiny topic pertaining to something we've learned together this semester and then submit a code snippet to be presented in class on Wednesday, 11/12. (Think \~2 minute walkthrough)

#### Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidyverse)
library(here)
library(janitor) 
library(arrow) # lets us read/write parquet and feather files
library(sf) # "simple features" for spatial data
library(urbnmapr) # access to US state/country shapefiles for mapping
library(naniar) # tools for exploring and visualizing missing data
library(ggiraph) # makes ggplot charts interactive
options(scipen = 99) # turns off scientific notation
```

#### Read in the list of topics

```{r, message=FALSE}
topics <- read_csv("data/cs201_topics.csv")
view(topics)

# anything to edit?
# topics[7, "topic"] <- "wildcard"
```

#### Read in the list of student names

```{r, message=FALSE}
names <- read_csv("data/cs201_names.csv")
view(names)
```

#### Generate a random number in the names dataframe

```{r, message=FALSE}
names <- names %>%
  mutate(number = sample(1:nrow(topics), nrow(names), replace = FALSE))

view(names)
```

#### Join the names to the topics

```{r, message=FALSE}
assigned_topics <- names %>%
  full_join(topics, by = "number")

print(assigned_topics)

write.csv(assigned_topics, "data/assignment9_assigned_topics.csv")

# note that if you run the above script and randomize the names again, your outcome will be different from mine!!
# go with the topics assigned in class and announced on Canvas!
```

### Instructions

1.  Find your assigned topic. Talk to me if you reeeeeeaaaally want to do something else, or trade with someone (but let me know).

2.  Prepare a code snippet that briefly (\~2) explains the topic in the code chunk below, using `penguins` data.

3.  Push to GitHub **BY 10:00am** on Wednesday 11/10 (I'll need time to compile before class).

4.  Come to class Wednesday and show us how it's done :)

#### Extra libraries needed (if any)

```{r, message=FALSE}
library(janitor)

```

#### Read in penguins data

```{r, message=FALSE}
penguins <- read_csv(here("data/penguins_lter.csv"))

penguins <- penguins %>% 
  clean_names()
```

#### {Insert topic here}

```{r, message=FALSE}
# insert code snippet here
ggplot(data = penguins, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot() +
  geom_jitter(size = 0.5) + # adding some datapoints and jittering them
  theme_bw() + # sets overall plot theme (think background, axes, bounding box)
  scale_fill_viridis_d() +
  labs(title = "Palmer Penguins, Distribution of Flipper Length",
       subtitle = "3 penguin species observed on three islands in the Palmer Archipelago, Antarctica",
       x = "",
       y = "flipper length (mm)") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") +
  coord_flip()

```

```{r}

```
